---
author: "Coleman Harris"
title: "An Introduction to Shiny"
date: 01/30/2020
output:
  ioslides_presentation:
      css: shiny-slides.css
      logo: images/shiny-logo.png
      self_contained: no
---
```{r include=FALSE}

```

## Background

### About Me

- PhD Student in the Department of Biostatistics
- Researching spatial analysis methods for cell imaging data
- Dedicated R user, occasional Python & SQL user
- More info, projects, links: http://colemanrharris.me

### Slides

- Using an `ioslides` presentation format with Rmarkdown
- [Source](https://github.com/rstudio/ShinyDeveloperConference/tree/master/Debugging) for the styles and Shiny logo

## Introducing Shiny

### The basics

- Shiny is an R package that lets you wrap R code into interactive web applications
- Useful for dashboards, interactive exploratory data analysis, and more (examples at the end)

### Resources

- The [Shiny Gallery](https://shiny.rstudio.com/gallery/)
- The official [tutorial](https://shiny.rstudio.com/tutorial/)
- Shiny [cheat sheet](https://shiny.rstudio.com/articles/cheatsheet.html)

## Structure

- Shiny apps are stored in one file, called `app.R`, and all related files live in a single directory
- Each Shiny app has 3 components:
    - **a user interface object** (`ui`)
        - This determines the structure and visual components of the app
    - **a server function** (`server`)
        - This code is how the app output interacts with the data and app inputs
    - **call to the** `shinyApp` **function**
        - Runs the application
    
## Structure

- Your app will look something like this:
```{r, eval = F}
library(shiny)

ui <- ... ## this sets up the visuals

server <- ... ## this sets up the app

shinyApp(ui = ui, server = server) ## this runs the app
```

# Old Faithful Example

## `ui`: The visual & input components

```{r eval=F}
ui = bootstrapPage(
    ## dropdown menu to select bin number
    selectInput(inputId = "n_breaks",
                label = "Number of bins in histogram (approximate):",
                choices = c(10, 20, 35, 50), selected = 20),
    ## checkbox to add density estimate to plot
    checkboxInput(inputId = "density",
                  label = strong("Show density estimate"),
                  value = FALSE),
    ## output call to setup the plot
    plotOutput(outputId = "main_plot", height = "300px"),
    # Display this only if the density is shown
    conditionalPanel(condition = "input.density == true",
                     sliderInput(inputId = "bw_adjust",
                                 label = "Bandwidth adjustment:",
                                 min = 0.2, max = 2, value = 1, 
                                 step = 0.2)
                     ))
```

## `server`: Generating the output
```{r eval=F}
server = function(input, output) {
    output$main_plot <- renderPlot({
      ## histogram of the eruptions
      hist(faithful$eruptions,
           probability = TRUE,
           breaks = as.numeric(input$n_breaks),
           xlab = "Duration (minutes)",
           main = "Geyser eruption duration")
      ## add density curve if input is TRUE
      if (input$density) {
        dens <- density(faithful$eruptions,
                        adjust = input$bw_adjust)
        lines(dens, col = "blue")
      }
    })
}
```

## The app!
```{r eval=F}
## load in shiny
library(shiny)

## load in the data
data("faithful")

## run your app
runApp(ui = ui, server=server)
```

- Now let's check it out in RStudio!
  - The **Run App** shortcut
  - https://colemanrharris.shinyapps.io/faithfulApp/

## Checking in--

### Now we can...
- Read & follow the structure of a Shiny app
- Incorporate different inputs and outputs into the app
- Run the Shiny app
- Publish a Shiny app on http://shinyapps.io

### Now what?

## The wonderful world of Shiny

- Reactive [example](https://colemanrharris.shinyapps.io/HinchmanForUSCongress-VoterApplication/)
- `shinydashboard` ([Example](https://edgarruiz.shinyapps.io/db-dashboard/))
- [Shiny gadgets](https://shiny.rstudio.com/articles/gadgets.html)
- Amazing examples:
    - [iSEE - interactive Summarized Experiment Explorer](https://kevinrue.shinyapps.io/isee-shiny-contest/)
    - [69 Love Songs: A Lyrical Analysis](https://committedtotape.shinyapps.io/sixtyninelovesongs/)
    - [rstudio::conf(2019) Tweet Dashboard](https://gadenbuie.shinyapps.io/tweet-conf-dash/)
    - [Pet Records](https://jennadallen.shinyapps.io/pet-records-app/)
    - ...and many [more](https://blog.rstudio.com/2019/04/05/first-shiny-contest-winners/)!